Crop adaptation through gene flow

### 2016-07-10 Crop_adaptation_gene_flow_gj_outline_shortened.tex
Update Crop_adaptation_gene_flow_gj_outline_shortened.tex

Maize is a classical model in crop species to study the issue of adaptive introgresssion from wild relatives, as it still grows in sympatry with many of its wild relatives \cite{hufford2013}.

Maize was domesticated about 10,000 years ago in Mexico \cite{smith1997initial} \cite{hufford2012comparative} in the Balsas River valley \cite{matsuoka2002single} from teosinte.

Mexican farmers have long been aware of maize-teosinte hybrids, and in some places, such as the Nobogame valley, farmers have observed desirable traits passed to their crops through the hybrid plants \cite{wilkes1977hybridization} \cite{lumholtz1902unknown} \cite{wilkes1970teosinte}.

In particular, \emph{Zea mays} ssp. \emph{mexicana} (referred to \emph{mexicana} hereafter) and Mexican highland maize have become a well-studied pair (although ssp. \emph{parvaglumis} has also been shown to hybridize with maize, \cite{wilkes1977hybridization}).

Introgression between \emph{mexicana} and Mexican highland maize has been reported based on evidence from both morphological data \cite{wilkes1977, lauter2004, doebley1984} and molecular analysis \cite{matsuoka2002, vanHeerwaarden2011, doebley1987, warburton2011, fukunaga2005}. [[Li's and my bib tags might be in conflict, e.g. wilkes1977 is likely my wilkes1977hybridization.]]
.
 Doebley (\cite{doebley1987patterns}) found that north-central high-elevation Mexican landraces (Apachito, Arrocillo Amarillo, Azul, Celaya, Chaloqueno, Conico, Conico Norteno, Gordo, Harinoso de Ocho Occidentales)[[special characters]] show evidence of introgression of two alleles from \emph{mexicana}, and suspected that they might have fitness advantages for adapting to highland environmental conditions.

Population structure analyses by \cite{matsuoka2002single} estimate that \emph{mexicana} gene flow has contributed to the gene pools of the Cacahuacintle, Palomero de Jalisco, and Palomero Taloqueno landraces, but also into (non-landrace) Mexican maize at the same elevation.

@@

\cite{bitocchi2009introgression} has a good review of the history of domestication, introduction to Europe, and further development.

>>>>

Maize is an ideal model system for the study of adaptive introgresssion from wild relatives.
Mexican farmers have long been aware of maize-teosinte hybrids, and in some places, such as the Nobogame valley, farmers have observed desirable traits passed to their crops through the hybrid plants \cite{wilkes1977hybridization} \cite{lumholtz1902unknown} \cite{wilkes1970teosinte}.
 Doebley (\cite{doebley1987patterns}) found that north-central high-elevation Mexican landraces (Apachito, Arrocillo Amarillo, Azul, Celaya, Chaloqueno, Conico, Conico Norteno, Gordo, Harinoso de Ocho Occidentales)[[special characters]] show evidence of introgression of two alleles from \emph{mexicana}, and suspected that they might have fitness advantages for adapting to highland environmental conditions.
 Population structure analyses by \cite{matsuoka2002single} estimate that \emph{mexicana} gene flow has contributed to the gene pools of the Cacahuacintle, Palomero de Jalisco, and Palomero Taloqueno landraces, but also into (non-landrace) Mexican maize at the same elevation.
In particular, \emph{Zea mays} ssp. \emph{mexicana} (referred to \emph{mexicana} hereafter) and Mexican highland maize have become a well-studied pair (although ssp. \emph{parvaglumis} has also been shown to hybridize with maize, \cite{wilkes1977hybridization}).
Introgression between \emph{mexicana} and Mexican highland maize has been reported based on evidence from both morphological data \cite{wilkes1977, lauter2004, doebley1984} and molecular analysis \cite{matsuoka2002, vanHeerwaarden2011, doebley1987, warburton2011, fukunaga2005}. [[Li's and my bib tags might be in conflict, e.g. wilkes1977 is likely my wilkes1977hybridization.]]

@@
